### 2019-03-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for now, I have to go for my teaching duties.

\TODOL{Be much more concise, if I already give an abstract in the beginning of the chapter...}

@@

\TODOL{Be much more concise: in the thesis, MAB models and applications have already been introduced, motivated etc!}

@@

Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy (\cite{Thompson33}). More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, or parameters of a dynamically configurable radio hardware (\cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}). Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
%
For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not evolve over time} may be a big limitation. Indeed, in cognitive radio new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate; whereas in online recommendation, the popularity of items is also subject to trends. Hence, there has been some interest on how to take those \emph{non-stationary} aspects into account within a multi-armed bandit model.

A first possibility to cope with non-stationary is to model the decision making problem as an \emph{adversarial bandit problem} (\cite{Auer02NonStochastic}). Under this model, rewards are completely arbitrary and are not assumed to follow any probability distribution.
For adversarial environments, the pseudo-regret, which compares the accumulated reward of a given strategy with that of the best fixed-arm policy, is often studied. The pseudo regret of the EXP3 algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}. However, this model is a bit too general for the considered applications, where reward distributions do not necessarily vary at every round. For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced by \cite{Kocsis06} and \cite{YuMannor09}.
% It is sometimes referred to as the abruptly changing MAB, or switching environments (\cite{MellorShapiro13}).
%
In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $i$ at round $t$ has some mean $\mu_i(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $i_t^* = \arg\max_{i} \mu_i(t)$.

In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB. Like previous approaches -- \CUSUM{} (\cite{LiuLeeShroff17}) and \MUCB{} (\cite{CaoZhenKvetonXie18}) -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector. For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 (\cite{Auer02}) used in previous works. For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest. This choice is particularly appealing because unlike previous approaches, the Bernoulli GLRT is \emph{parameter-free}: it does not need the tuning of a window size ($w$ in \MUCB), or the knowledge of a lower bound on the magnitude of the smallest change ($\varepsilon$ in \CUSUM).

In this work we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
Our numerical simulations reveal that using local restart leads to better empirical performance, and show that our approach often outperforms existing competitors.

\paragraph{Outline}
%
This chapter is structured as follows.
We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}. In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test (GLRT) as a Change-Point Detector (CPD) algorithm.
We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bound on the regret of each variant. The unified regret analysis for these two algorithms is sketched in Section~\ref{sec:6:RegretAnalysis}. Numerical experiments
are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.

@@

In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
We define the $k$-th breakpoint by
\begin{equation}
    \tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}.
\end{equation}
%
Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, the rewards $(X_{i,t})$ associated to each arms are \iid.

Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed. Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only a few arms change at each breakpoint. Introducing the number of change-points  on arm $i$, defined as $\NCi:=\sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
$\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms. Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.

\paragraph{An interesting generalization}
%
The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while the adversarial model can be considered as a special (worst) case, with $\Upsilon_T = T$. However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.

@@

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To the authors knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitors the rewards of each arm by using a change detection algorithm (\cite{Basseville93}), and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies are tuned, with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.

@@

Both D-UCB and SW-UCB builds on a stationary policy,
but for example EXP3.S from \cite{Auer02NonStochastic}
builds on the EXP3 policy, which is designed for the adversarial case,
and achieves a larger regret upper-bound.
We also found in our experiments that it always performed worse that the reference stationary algorithms (\eg, \klUCB), so we do not include it in Section~\ref{sec:6:NumericalExperiments}.
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proved to be less efficient empirically, or are designed for more specific settings.
%
For instance, Windowed-Mean Shift from \cite{YuMannor09} is generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting as they consider side observations.
%
A different setting where the quantity of interest is not $\Upsilon_T$ but $V_T$ the total variational budget, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
%
The RExp3 algorithm (\cite{Besbes14stochastic}) can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm. Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point). A  $\cO(V_T^{1/3} T^{2/3})$ regret bound is proved, which is weaker than existing results in our setting. Hence we do no include this algorithm in our experiments.
%

@@

\paragraph{Actively adaptive algorithms.}
%
We distinguish two families,
the first line of research uses frequentist change-point detectors (\cite{Basseville93}) combined with stationary policies, usually index policies like \UCB.
When using an efficient CD algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley Test and the \UCB{} policy, but no theoretical guarantee was given.
The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a a suboptimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proved, even when $\Upsilon_T$ is known.

@@

Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametrized by $\alpha\in(0,1)$.
\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

@@

\paragraph{Advantages of our approach}
%
\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
This study is supported by both theoretical and empirical results.
%
Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing  much in terms of regret.

\paragraph{About \klUCB.}
%
Our proposal \GLRklUCB{} is inspired by both the \MUCB{} and \CUSUMUCB{} algorithms.
Previous works focussed on using \UCB{} (\cite{LiuLeeShroff17,CaoZhenKvetonXie18}),
but we propose to use \klUCB{} instead, as it is known to be more efficient for Bernoulli rewards as well as for a more generic case of bounded or one-dimensional exponential families (\cite{KLUCBJournal}).
Theoretical results for these works were only given for \UCB, but they both suggest that extending the results to \klUCB{} (or other efficient stationary policies) should not be difficult.
For a fair comparison, we therefore chose to compare the different Change-Point detector algorithms combined with \klUCB.

% -----------------------------------------------------------------
\section{The Bernoulli GLR Change-Point Detector}
\label{sec:6:ChangePointDetector}

Sequential change-point detection has been extensively studied in the statistical community (see, \eg, \cite{Basseville93} for a survey). In this section, we are interested in detecting changes on the mean of a probability distribution with bounded support. Assuming that  we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$. We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau \in\N^*$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a stopping time $\hat\tau$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N, \bE[X_i] = \mu_0\right)$.

@@

\TODOL{We need to include and prove our results for false alarm, delay, etc, all results we squeezed and compressed for the COLT version \cite{Besson2019GLRT}.}

@@

If the samples $(X_t)$ were all drawn from a Bernoulli distribution, our change-point detection problem would reduce to a parametric sequential test of
$\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$
against the alternative
$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \ldots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \ldots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
The Generalized Likelihood Ratio statistic for this test is defined by
\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < t}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_t ; \mu_0)},\]
%
where $\ell(X_1, \ldots, X_t ; \mu_0)$ and $\ell(X_1, \ldots, X_t ; \mu_0,\mu_1,\tau)$ denote the likelihoods of the first $t$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$.
Using the form of the likelihood for Bernoulli distribution, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
%
\begin{equation}\label{eq:6:BernoulliDivergence}
    \kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right).
\end{equation}
%
Indeed, one can show that $\GLR(t) = \sup_{s \in [1,t]} \left[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:t}\right) + (t-s) \times \kl\left(\hat{\mu}_{s+1:t},\hat{\mu}_{1:t}\right)\right]$,
where for $k \leq k'$, $\hat{\mu}_{k:k'}$ denotes the average of the observations collected between the instants $k$ and $k'$. This motivates the definition of the Bernoulli GLR change point detector.

\begin{definition}\label{def:6:GLRDef}
    The Bernoulli GLR change point detector with threshold function $\beta(n,\delta)$ is
    %
    \begin{equation}\label{def:6:GLR}
        \hat\tau_{\delta} := \inf \left\{ n \in \N^* : \sup_{s \in [1,t]} \left[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
    \end{equation}
\end{definition}

Asymptotic properties of the GLR for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR} but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence between two elements in that exponential family that have mean $x$ and $y$. For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to \eqref{def:6:GLR} with $\kl(x,y) = 2(x-y)^2$ when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.

In this section, we provide new non-asymptotic results about the Bernoulli GLR test under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the Bernoulli GLR may stop earlier that the Gaussian GLR, based on the quadratic divergence $2(x-y)^2$.

% ----------------------------------------------------------------------------
\subsection{Non-Asymptotics Properties of the Bernoulli GLR}\label{subsec:6:PropGLR}

When used for a bandit problem, the two main properties of a sequential change-point detection test are its false alarm probability and its detection delay.
%
A small false alarm probability ensures that no detection occurs before it should (\ie, no useless detection occur on stationary segments), and a small detection delay ensures that, if the stationary segments are long enough, then every change-points on every arms will be detected after a short enough amount of samples from that arm.
%
We give below two Lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:DetectionDelay} which bounds the detection delay if there is enough samples before a change.

%-----------------------------------------------------------------------------
\paragraph{False alarm probability}

>>>>

% \TODOL{Be much more concise, if I already give an abstract in the beginning of the chapter...}

@@

% \TODOL{Be much more concise: in the thesis, MAB models and applications have already been introduced, motivated etc!}

@@

Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
%
For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not evolve over time} may be a big limitation. Indeed, in cognitive radio new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate; whereas in online recommendation, the popularity of items is also subject to trends. Hence, there has been some interest on how to take those \emph{non-stationary} aspects into account within a multi-armed bandit model.

A first possibility to cope with non-stationary is to model the decision making problem as an \emph{adversarial bandit problem} \cite{Auer02NonStochastic}. Under this model, rewards are completely arbitrary and are not assumed to follow any probability distribution.
For adversarial environments, the pseudo-regret, which compares the accumulated reward of a given strategy with that of the best fixed-arm policy, is often studied. The pseudo regret of the EXP3 algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}. However, this model is a bit too general for the considered applications, where reward distributions do not necessarily vary at every round. For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced by \cite{Kocsis06} and \cite{YuMannor09}.
It is sometimes referred to as the abruptly changing MAB, or switching environments \cite{MellorShapiro13}.
%
In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $i$ at round $t$ has some mean $\mu_i(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $i_t^* = \arg\max_{i} \mu_i(t)$.

In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB. Like previous approaches -- \CUSUM{} \cite{LiuLeeShroff17} and \MUCB{} \cite{CaoZhenKvetonXie18} -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector. For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 \cite{Auer02} used in previous works. For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest. This choice is particularly appealing because unlike previous approaches, the Bernoulli GLRT is \emph{parameter-free}: it does not need the tuning of a window size ($w$ in \MUCB), or the knowledge of a lower bound on the magnitude of the smallest change ($\varepsilon$ in \CUSUM).

In this chapter we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
Our numerical simulations reveal that using local restart leads to better empirical performance, and show that our approach often outperforms existing competitors.

\paragraph{Outline.}
%
This chapter is structured as follows.
We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}. In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a Change-Point Detector (CPD) algorithm.
We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on the regret of both variants.
The unified regret analysis for these two algorithms is given in Section~\ref{sec:6:RegretAnalysis}.
Numerical experiments are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.

@@

In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint by
\begin{equation}
    \tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}.
\end{equation}
%
Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, that is on any stationary segments, the rewards $(X_{i,t})$ associated to each arms are \iid{} (hence the name of \emph{piece-wise stationary} problems).

Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fix channel or subset of channels, and on a particular day in a particular city, if the company deploys a lot of new objects, then one or some channels see their mean occupancy rates abruptly change.

Introducing the number of change-points on arm $i$, defined as $\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
$\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms. Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.

\paragraph{An interesting interpolation.}
%
The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while the adversarial model can be considered as a special (worst) case, with $\Upsilon_T = T$. However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.

@@

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To the authors knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitors the rewards of each arm by using a change detection algorithm \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies are tuned, with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.

@@

Both D-UCB and SW-UCB builds on a stationary policy,
but for example EXP3.S from \cite{Auer02NonStochastic}
builds on the EXP3 policy, which is designed for the adversarial case,
and achieves a larger regret upper-bound.
We also found in our experiments that it always performed worse that the reference stationary algorithms (\eg, \klUCB), so we do not include it in Section~\ref{sec:6:NumericalExperiments}.
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proved to be less efficient empirically, or are designed for more specific settings.
%
For instance, Windowed-Mean Shift from \cite{YuMannor09} is generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting as they consider side observations.
%
A different setting where the quantity of interest is not $\Upsilon_T$ but $V_T$ the total variational budget, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
%
It can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm. Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point). A  $\cO(V_T^{1/3} T^{2/3})$ regret bound is proved, which is weaker than existing results in our setting. Hence we do no include this algorithm in our experiments.
%

@@

\paragraph{Actively adaptive algorithms.}
%
We distinguish two families,
the first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
When using an efficient CD algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley Test and the \UCB{} policy, but no theoretical guarantee was given.
The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a a suboptimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proved, even when $\Upsilon_T$ is known.

@@

Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

@@

\paragraph{Advantages of our approach.}
%
\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
This study is supported by both theoretical and empirical results.
%
Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing  much in terms of regret.

\paragraph{About \klUCB.}
%
Our proposal \GLRklUCB{} is inspired by both the \MUCB{} and \CUSUMUCB{} algorithms.
Previous works focussed on using \UCB{} \cite{LiuLeeShroff17,CaoZhenKvetonXie18},
but we propose to use \klUCB{} instead, as it is known to be more efficient for Bernoulli rewards as well as for a more generic case of bounded or one-dimensional exponential families \cite{KLUCBJournal}.
Theoretical results for these works were only given for \UCB, but they both suggest that extending the results to \klUCB{} (or other efficient stationary policies) should not be difficult.
For a fair comparison, we therefore chose to compare the different change-point detector algorithms combined with \klUCB.

% -----------------------------------------------------------------
\section{The Bernoulli GLRT Change-Point Detector}
\label{sec:6:ChangePointDetector}

Sequential change-point detection has been extensively studied in the statistical community (see, \eg, \cite{Basseville93} for a survey). In this section, we are interested in detecting changes on the mean of a probability distribution with bounded support. Assuming that  we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$. We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a stopping time $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N^*, \bE[X_i] = \mu_0\right)$.

@@

\TODOL{We need to PROVE our results for false alarm and delay !.}

@@

If the samples $(X_t)$ were all drawn from a Bernoulli distribution, our change-point detection problem would reduce to a parametric sequential test of
\[\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0)),\]
against the alternative
\[\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \ldots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \ldots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1)).\]

The Generalized Likelihood Ratio statistic for this test is defined by
\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},\]
%
where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$ over $\cH_1$.
%
By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
%
\begin{equation}\label{eq:6:BernoulliDivergence}
    \kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right).
\end{equation}
%
Indeed, one can show that $\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right]$,
where for $k \leq k'$, $\widehat{\mu}_{k:k'} = \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
%
This motivates the following definition of the Bernoulli GLRT change point detector.

\begin{definition}\label{def:6:GLRDef}
    The Bernoulli GLRT (B-GLRT) change point detector with threshold function $\beta(n,\delta)$ is the stopping time $\widehat{\tau}_{\delta}$ defined by
    %
    \begin{equation}\label{def:6:GLR}
        \widehat{\tau}_{\delta} := \inf \left\{ n \in \N^* : \sup_{s \in [1,t]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
    \end{equation}
\end{definition}

Asymptotic properties of the GLRT for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR} but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence between two elements in that exponential family that have mean $x$ and $y$. For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to \eqref{def:6:GLR} with $\kl(x,y) = 2(x-y)^2$ when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.

In this section, we provide new non-asymptotic results about the B-GLRT under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $2(x-y)^2$.

% ----------------------------------------------------------------------------
\subsection{Non-Asymptotics Properties of the B-GLRT}\label{subsec:6:PropGLR}

When used for a bandit problem, the two main properties of a sequential change-point detection test are its false alarm probability and its detection delay.
%
A small false alarm probability ensures that no detection occurs before it should (\ie, no useless detection occur on stationary segments), and a small detection delay ensures that, if the stationary segments are long enough, then every change-points on every arms will be detected after a short enough amount of samples from that arm.
%
We give below two Lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there is enough samples before a change.

%-----------------------------------------------------------------------------
\paragraph{False alarm probability.}
